A Study of Residual Adapters for Multi-Domain Neural Machine Translation

### 2020-10-06 2020/wmt_20/mdadapt.tex
local corrections following reviews

\newcommand{\mpTodo}[1]{\Todo[MP:]{\textcolor{green}{#1}}}
\newcommand{\mpDone}[1]{\done[MP]\Todo[MP:]{\textcolor{green}{#1}}}
\usepackage{mathtools,xparse}
\DeclarePairedDelimiter{\abs}{\lvert}{\rvert}
\DeclarePairedDelimiter{\norm}{\lVert}{\rVert}
\NewDocumentCommand{\normL}{ s O{} m }{%
  \IfBooleanTF{#1}{\norm*{#3}}{\norm[#2]{#3}}_{2}%
}
\newcommand{\src}{\ensuremath{\mathbf{f}}} % source sentence
\newcommand{\trg}{\ensuremath{\mathbf{e}}} % target sentence
\newcommand{\domain}[1]{\texttt{\textsc{#1}}}
\newcommand{\system}[1]{\texttt{\textbf{#1}}}
\newcommand{\vlambda}{\ensuremath{\boldsymbol\lambda}\xspace} % parameters vector for a distribution
\newcommand{\indic}[1]{\ensuremath{\mathbb{I}(#1)}}
% \newcommand{\SB}[1]{\textcolor{green}{#1}}
% \newcommand{\SW}[1]{\textcolor{red}{#1}}
\newcommand{\SB}[1]{\textbf{#1}}
\newcommand{\SW}[1]{\underline{#1}}
\renewcommand\textfraction{.1}
\renewcommand\floatpagefraction{.8}

@@

\date{}

\begin{document}
\maketitle
\begin{abstract}
% Among the approaches for adapting a pretrained NMT model to a specific domain, finetuning is the dominant method. Finetuning proposes 2 approaches: simply continuing training all the parameters \cite{Luong15stanford}; training addition adapters while freezing the pretrained model\cite{Bapna19simple,Vilar18learning}. The second approach has several advantages including preserving the pretrained model and the legerity of the adapters. However, the behavior of these adapters has not been further studied. The objective of this paper is to give an ablation study on the use of residual adapter in the domain adaptation problem.
% \mpTodo{correcting abstract}
% **** 
\fyDone{Citation-free abstract}
Domain adaptation is an old and vexing problem for machine translation systems. The most common and successful approach to supervised adaptation is to fine-tune a baseline system with in-domain parallel data. Standard fine-tuning however modifies all the network parameters, which makes this approach computationally costly and prone to overfitting. A recent, lightweight approach, instead augments a baseline model with supplementary (small) adapter layers, keeping the rest of the model unchanged. This has the additional merit to leave the baseline model intact and adaptable to multiple domains. In this paper, we conduct a thorough analysis of the adapter model in the context of a multidomain machine translation task. We contrast multiple implementations of this idea using two language pairs. Our main conclusions are that residual adapters provide a fast and cheap method for supervised multi-domain adaptation; our two variants prove as effective as the original adapter model and open perspective to also make adapted models more robust to label domain errors.
\fyDone{abstract to be continued}

\end{abstract}
\section{Introduction \label{sec:intro}}
\mpDone{write introduction} \fyDone{Citations in chronological order}\fyDone{Split long sentences}
Owing to multiple improvements, Neural Machine Translation (NMT) \cite{Kalchbrenner13recurrent,Sutskever14sequence,Bahdanau15learning,Vaswani17attention} nowadays delivers useful outputs for many language pairs. However, as many deep learning models, NMT systems need to be trained with sufficiently large amounts of data to reach their best performance. Therefore, the quality of translation of NMT models is still limited in low-resource language or domain conditions \cite{duh13adaptation,zoph16transfer,koehn17six}. While many approaches have been proposed to improve the quality of NMT models in low-resource domains (see the recent survey of \citet{Chu18asurvey}), full fine-tuning \cite{Luong15stanford,neubig18rapid} of a generic baseline model remains the dominant supervised approach when adapting NMT models to specific domains.

Under this view, building adapted systems is a two-step process: (a) one first trains NMT with the largest possible parallel corpora, possibility aggregating texts from multiple, heterogeneous sources; (b) assuming that in-domain parallel documents are available for the domain of interest, one then adapts the pre-trained model by resuming training with the sole in-domain corpus. It is a conjecture that pre-trained model constitutes a better initialization than a random one, especially when adaptation data is scarce. Indeed, studies of transfer learning for NMT such as \cite{artetxe20cross,aji20neural} have confirmed this claim in extensive experiments. Full fine-tuning, that adapts all the parameters of a baseline model usually significantly improves the quality of the NMT for the chosen domain. However, it also yields large losses in translation quality for other domains, a phenomenon referred to as ``catastrophic forgetting'' in the neural network literature \cite{McCloskey89catastrophic}. Therefore, a fully fine-tuned model is \emph{only useful to one target domain}. As the number of domains to handle grows, training and maintaining a separate model for each task can quickly become tedious and resource-expensive .\fyDone{Fix this sentence.}

\cite{Vilar18learning,Bapna19simple} have recently proposed a simple and lightweight domain adaptation method that also preserves the value of pre-trained models, based on small adapters component that are plugged in each hidden layer. These adapters are trained only with the in-domain data, keeping the pre-trained model frozen. Because these additional adapters are very small compared to the size of the baseline model, their use significantly reduces the cost of training and maintaining fine-tuned models, while still delivering a performance that is close to that of full fine-tuning.

In this paper, we would like to extend this architecture to improve NMT in several settings that still challenge automatic translation, such as translating texts from multiple topics, genre or domains, in the face of unbalanced data distributions. Furthermore, as the notion of ``domains''  is not always well established, another practical setting is the translation of texts mixing several topics / domains. An additional requirement is to translate texts from domains unseen in training, based only on the unadapted system, which should then be made as strong as possible. 
% Residual adapters allow us to adapt NMT model to any specific domain in a computationally cheap way. Could we adapt NMT model to noisy text and topical text at once? While residual adapters are good at adapting to one specific domain, their performance still dramatically decrease for the other domains seen in training, as we show in our experiments. Therefore, we have to decide whether to use residual adapters manually, i.e, we have to know to which domain the text belongs a priori. Therefore, we would like to fuse a domain classifier to the architecture in order to weight the contribution of the adapters with respect to the topic-relatedness of the text.
\fyDone{Say differently: various implementations}

@@

\section{Residual adapters \label{sec:res}}
In this section, we describe the basic version of the residual adapter architectures, as well as two original variants of this model.

@@

\subsubsection{The computation of adapter layers}
Our reference architecture is the Transformer model of \cite{Vaswani17attention}, which we assume contains a stack of layers both on the encoder and the decoder sides. Each layer contains two subparts, an attention layer and a dense layer. Details vary from one implementation to another, we simply contend here that each layer $i \in \{1 \dots L\}$ (in the encoder or the decoder) computes a transform of a fixed-length sequence of $d$-dimensional input vectors $h^{i}$ into a sequence of output vectors $h^{i+1}$.

The residual adapter architecture extends this model by adding one new computation module for each layer. The $i^{\text{th}}$ residual adapter thus modifies $h^i$ by computing the following transformations:\fyDone{Use align env}
\begin{align*}
  h^{i}_1 &= \mathbf{W}_{db}^{i}h^{i} + b^i_{1} \\
  h^{i}_2 &= \mathbf{ReLU}(h_1^{i})
  \\
%\end{align*}
%\begin{align*}
  h^{i}_3 &= \mathbf{W}_{bd}^{i}h_2^{i} + b^i_{2} \\
  \bar{h}^{i} &= h^{i}_3 + h^i.
\end{align*}
Overall, the  $i^{\text{th}}$ adapter is thus parameterized by matrices $\displaystyle{\mathbf{W}_{db}^{i}\in\mathbb{R}^{d\times b}}$,$\displaystyle{\mathbf{W}_{bd}^{i}\in\mathbb{R}^{b\times d}}$, bias vectors $\displaystyle{b^i_{1} \in \mathbb{R}^{b}}$, $\displaystyle{b^i_{2} \in \mathbb{R}^{d}}$, with $b$ the dimension of the adapter ($d \gg b$)\fyDone{Check this}. For the sake of brevity, we will simply denote $h^{i}_3 = \operatorname{ADAP}^{(i)}(h^i)$, and $\theta_{\operatorname{ADAP}^{(i)}}$ the corresponding set of parameters.\fyDone{or is it $h_i$ ?}\fyDone{attention aux matrices $W_i$}

@@

\subsubsection{Design space and variants \label{sssec:design-space}}
This general architecture leaves open many design choices pertaining to the details of the network organization, the training procedure and the corresponding objective function.

A first question concerns the number of adapter layers. While in principle, all Transformer layers can be subject to adaptation, it is nonetheless worthwhile to consider simpler adaptation schemes, which would only alter a limited number of layers. Such strategy might be especially relevant when the training data contains very small domains, as in the experiments of Section~\ref{sec:exp}, and for which a complete adaptation may not be necessary or/and or prone to overfitting. Likewise, it might be meaningful to explore ways to share subsets of adapters across domains. This in turn raises the issue of which layer(s) to adapt, a question that can be approached in the light of recent analyses of Transformers models, which conjecture that the higher layers encode global patterns with a more ``semantic'' interpretation, while the lower layers encode local patterns akin to morpho-syntactic information \cite{raganato18analysis}.

% The size of available corpora for each domain is extremely varying. For example, the corpus of domain \domain{tourism} in En-De contains only around $7\times10^3$ sentence pairs while the corpus of domain \domain{news} in En-De contains approximately $3\times10^6$ sentence pairs. For extremely small domains, we could economize the number of parameters by reducing the number of residual adapters in the architecture. While only a limited number of adapters are incorporated to the NMT model, which positions in the model are more important in the domain adaptation task? We would like to analyze the impact of position and number of residual adapters involved in the adapted model. It is conjectured that the higher layers represent more global patterns such as semantic while the lower layers represent more local patterns such as syntactic \cite{raganato18analysis}\fyTodo{Missing reference here}. The domain shift in local patterns and global patterns has not yet studied. In this paper, we do not intend to study this aspect of the domain adaptation problem. We would like to give an extensive comparison between domain adaptation in different levels in NMT model.

A related question concerns the regularization of adapter layers to mitigate overfitting. Reducing the number of adapters, or their dimensions, is simple, but such choices are difficult to optimize numerically -- an issue that becomes important as the number of domain grows. Less naïve alternatives can also be entertained, such as applying weight decay or layer regularization to the adapter. Implementing these requires to modify the objective function in a way that still allows for a smooth optimization problem. For instance, weight decay applies a penalization on the weights of the adapters, complementing the cross-entropy term with a function of the norm of the parameters: \fyDone{The second summation also runs over $x,y$ ? I think not}
\begin{equation*}
  \begin{split}
    \bar{L} & = \frac{1}{\#(x,y)}\mathop{\sum}_{x,y}( - \log(P(y|x))) \\
    & + \lambda  \sum_{i \in \{1,..,6\} \otimes \{enc, dec\}} \normL{\theta_{\operatorname{ADAP}^{(i)}}}
  \end{split}
\end{equation*}
An alternative scheme is \emph{layer regularization}, which penalizes the output of the adapters, corresponding to the following objective:
\begin{equation*}
  \begin{split}
    \bar{L} & = \frac{1}{\#(x,y)}\mathop{\sum}_{x,y} (-\log(P(y|x)) \\
    & + \lambda \sum_{i \in \{1,..,6\} \otimes \{enc, dec\}} \normL{\operatorname{ADAP}^{(i)}(h_i(x,y))})
  \end{split}
\end{equation*}

Finally, another independant design choice relates to the training strategy for adapters. A first option is to generalize supervised domain adaptation to multi-domain adaptation and to proceed in two steps: (a) train a generic model with all the available data; (b) train each adapter layer with domain-specific data, keeping the generic model parameters unchanged. Another strategy is to adopt the view of \citet{Dredze08online}, where the multi-domain setting is viewed as an instance of multi-task learning \cite{Caruana97multitask} with each domain corresponding to a specific task. This suggest to train all the parameters from scratch, as we would do in a multi-task mode. The generic parameters will still depend on all the available data, while each adapter will only be trained with the corresponding in-domain data.\fyFuture{Does everyone need a domain? Do we need this discussion - talk about sentence level adaptation, better classifier pour MDL}

@@

Therefore, we would like to design a version of residual adapters that is more robust to such domain errors. This variant, called the \emph{gated residual adapter model}, relies on the training of a supplementary trained component that will help decide whether to activate, on a word per word basis, a given residual layer, and to regulate the strength of this activation. To this end, we extend the highway version of residual adapters as follows.
\fyDone{Consistency of notations wrt section 2.1}

Formally, we replace the adapter computation of equation~\eqref{eq:highway-output} and take the adapted hidden (topmost) layer to be computed as (this is for domain $k$):
\begin{equation}
  \bar{h}^L = h^L + \displaystyle{\mathop{\sum}_{1 \leq i \leq L} \operatorname{ADAP}_k^i(h^i) \odot{} z_k(h^L)}, \label{eq:gated-output}
\end{equation}
where $z_k(h^L[t]) \in [0,1]$ measures the relatedness of the $t^{\text{th}}$ word $w_t$ to domain $k$. The more likely $w_t$ is in domain $k$, the larger $z_k(h^L[t])$ should be; conversely, for words\footnote{The term ``word'' is employed here by mere convenience, as systems only manipulate sub-lexical BPE units; furthermore, the values of the hidden representations $h^{i}$ at position $t$ depend upon all the other positions in the sentence.} that are not typical of any domain $k$ (eg.\ function words),  adaptation is minimum and the corresponding adapted encoder output ($\bar{h}^L[t]$) will remain close to the output of the generic model ($h^L[t]$). In our implementation, we incorporate two domain classifiers on top of the encoder and the decoder, that take the last hidden layer of the encoder (resp.\ decoder) as input and use the posterior probability $P(k|h^L[t])$ of domain $k$ as the value for $z_k(h^L[t])$.

Training gated residual adapters thus comprises three steps, instead of two for the baseline version:
\begin{enumerate}
\item learn a generic model with mixed corpora from multiple domains.
\item train a domain classifier on top of the encoder and decoder; during this step, the parameters of the generic model are frozen. This model computes the posterior domain probability $P(k|h^L[t])$ for each word $w_t$, based on the representation computed by the last layer.
\item train the parameters of adapters with in-domain data separately for each domain, while freezing all the other parameters.
\end{enumerate}
\fyFuture{is this classifier important, can we train with the rest of the system ? Yes we could, but do we want the encoder to be good at classification?}\fyFuture{Check writing to indicate label smoothing for this has been used.}\fyFuture{Can we train a more complex training model ? should we use more monolingual data ? etc}

@@

For all En$\rightarrow$Fr models, we set the embeddings size and the hidden layers size to~512. Transformers use multi-head attention with 8 heads in each of the 6 layers; the inner feedforward layer contains 2,048 cells. Residual adapters additionally use an adaptation block in each layer, composed of a 2-layer perceptron, with an inner ReLU activation function operating on normalized entries of dimension $b=1024$.
% The gated variant is made of a dense layer, followed by a layer normalization and a sigmoid activation.
% The domain control systems are exactly as their baseline counterparts (RNN and Transformer), with an additional 2 cells encoding the domain on the input layer.
Training use a batch size of~12,288 tokens; optimization uses Adam with parameters $\beta_1=0.9$, $\beta_2= 0.98$ and Noam decay ($warmup\_steps=4,000$), and a dropout rate of $0.1$ for all layers.\fyDone{Describe the block adaptation layer - voir slides}

@@

\iffalse{
\begin{table*}[htbp]
  \centering
  \begin{tabular}{|l|l|} \hline
    Model & params \\ \hline 
    \system{Transformer-En-Fr}  & 65 \\
    \system{Residual Adapter-En-Fr} & 1 \\
    \system{Transformer-En-De}  & 213 \\
    \system{Residual Adapter-En-De} & 8 \\
     \hline
  \end{tabular}
  \caption{Number of parameters ($\times 10^6$)}
  \label{tab:params}
\end{table*}
}
\fi

@@

By varying the number and position of residual adapters (see Section~\ref{ssec:architecture}), we then contrast several implementation of residual adapters. \fyDone{Fix style here} Because the set of possible configurations is large, we only perform experiments for layers $i= 2, 4, 6$ (both for the encoder and decoder). Two settings are considered: keeping just one adapter, or keeping the three. The trend is the same for the two language directions: suppressing adapters always hurts the overall performance, albeit by a small margin: having six adapters is better than three, which is better than keeping only one. With only one adapter active, we observe small, unsignificant changes in performance when varying the adapter's depth. Based on these experiments, it seems that adaptation remains useful even for the deeper layers of the architecture.\fyFuture{More configurations ?}

@@

Results in Tables~\ref{tab:performance-en-fr-pos-reg} and \ref{tab:performance-en-de-pos-reg} show that regularizing the adapter model can positively affect the test performance for the smallest domains (this is especially clear for weight-decay (\system{Res-Adap-WD}) in En$\rightarrow$De), at the cost however of a small drop in performance for the other domains. Using layer regularization proves here to be comparatively less effective. Finding better ways to set the regularization parameters, for instance by varying $\lambda$ for each domain based on the available supervision data, is left for future work.  
\fyDone{How is the weight decay parameter set ?}
\fyDone{Why is regularization not helping ? It helps for small domain - domain-specific regularization ??}

>>>>

\newcommand{\mpTodo}[1]{\Todo[MP:]{\textcolor{green}{#1}}}
\newcommand{\mpDone}[1]{\done[MP]\Todo[MP:]{\textcolor{green}{#1}}}
\usepackage{mathtools,xparse}
\DeclarePairedDelimiter{\abs}{\lvert}{\rvert}
\DeclarePairedDelimiter{\norm}{\lVert}{\rVert}
\NewDocumentCommand{\normL}{ s O{} m }{%
  \IfBooleanTF{#1}{\norm*{#3}}{\norm[#2]{#3}}_{2}%
}
\newcommand{\src}{\ensuremath{\mathbf{f}}} % source sentence
\newcommand{\trg}{\ensuremath{\mathbf{e}}} % target sentence
\newcommand{\domain}[1]{\texttt{\textsc{#1}}}
\newcommand{\system}[1]{\texttt{\textbf{#1}}}
\newcommand{\revision}[1]{\textcolor{red}{#1}}
\newcommand{\vlambda}{\ensuremath{\boldsymbol\lambda}\xspace} % parameters vector for a distribution
\newcommand{\indic}[1]{\ensuremath{\mathbb{I}(#1)}}
% \newcommand{\SB}[1]{\textcolor{green}{#1}}
% \newcommand{\SW}[1]{\textcolor{red}{#1}}
\newcommand{\SB}[1]{\textbf{#1}}
\newcommand{\SW}[1]{\underline{#1}}
\renewcommand\textfraction{.1}
\renewcommand\floatpagefraction{.8}

@@

\date{}
%\hypersetup{draft}
\begin{document}
\maketitle
\begin{abstract}
% Among the approaches for adapting a pretrained NMT model to a specific domain, finetuning is the dominant method. Finetuning proposes 2 approaches: simply continuing training all the parameters \cite{Luong15stanford}; training addition adapters while freezing the pretrained model\cite{Bapna19simple,Vilar18learning}. The second approach has several advantages including preserving the pretrained model and the legerity of the adapters. However, the behavior of these adapters has not been further studied. The objective of this paper is to give an ablation study on the use of residual adapter in the domain adaptation problem.
% \mpTodo{correcting abstract}
% **** 
\fyDone{Citation-free abstract}
Domain adaptation is an old and vexing problem for machine translation systems. The most common and successful approach to supervised adaptation is to fine-tune a baseline system with in-domain parallel data. Standard fine-tuning however modifies all the network parameters, which makes this approach computationally costly and prone to overfitting. A recent, lightweight approach, instead augments a baseline model with supplementary (small) adapter layers, keeping the rest of the model unchanged. This has the additional merit to leave the baseline model intact and adaptable to multiple domains. In this paper, we conduct a thorough analysis of the adapter model in the context of a multidomain machine translation task. We contrast multiple implementations of this idea using two language pairs. Our main conclusions are that residual adapters provide a fast and cheap method for supervised multi-domain adaptation; our two variants prove as effective as the original adapter model and open perspective to also make adapted models more robust to label domain errors.
\fyDone{abstract to be continued}

\end{abstract}
\section{Introduction \label{sec:intro}}
\mpDone{write introduction} \fyDone{Citations in chronological order}\fyDone{Split long sentences}
Owing to multiple improvements, Neural Machine Translation (NMT) \cite{Kalchbrenner13recurrent,Sutskever14sequence,Bahdanau15learning,Vaswani17attention} nowadays delivers useful outputs for many language pairs. However, as many deep learning models, NMT systems need to be trained with sufficiently large amounts of data to reach their best performance. Therefore, the quality of the translation of NMT models is still limited in low-resource language or domain conditions \cite{duh13adaptation,zoph16transfer,koehn17six}. While many approaches have been proposed to improve the quality of NMT models in low-resource domains (see the recent survey of \citet{Chu18asurvey}), full fine-tuning \cite{Luong15stanford,neubig18rapid} of a generic baseline model remains the dominant supervised approach when adapting NMT models to specific domains.

Under this view, building adapted systems is a two-step process: (a) one first trains NMT with the largest possible parallel corpora, possibility aggregating texts from multiple, heterogeneous sources; (b) assuming that in-domain parallel documents are available for the domain of interest, one then adapts the pre-trained model by resuming training with the sole in-domain corpus. It is a conjecture that the pretrained model constitutes a better initialization than a random one, especially when adaptation data is scarce. Indeed, studies of transfer learning for NMT such as \citet{artetxe20cross,aji20neural} have confirmed this claim in extensive experiments. Full fine-tuning, that adapts all the parameters of a baseline model usually significantly improves the quality of the NMT for the chosen domain. However, it also yields large losses in translation quality for other domains, a phenomenon referred to as ``catastrophic forgetting'' in the neural network literature \cite{McCloskey89catastrophic}. Therefore, a fully fine-tuned model is \emph{only useful to one target domain}. As the number of domains to handle grows, training, and maintaining a separate model for each task can quickly become tedious and resource-expensive .\fyDone{Fix this sentence.}

\citet{Vilar18learning,Bapna19simple,wuebker18compact,michel18extreme} have recently proposed a simple and lightweight domain adaptation method that also preserves the value of pre-trained models, based on small adapter components that are plugged in each hidden layer. These adapters are trained only with the in-domain data, keeping the pre-trained model frozen. Because these additional adapters are very small compared to the size of the baseline model, their use significantly reduces the cost of training and maintaining fine-tuned models, while still delivering a performance that is close to that of full fine-tuning.

In this paper, we would like to extend this architecture to improve NMT in several settings that still challenge automatic translation, such as translating texts from multiple topics, genre, or domains, in the face of unbalanced data distributions. Furthermore, as the notion of ``domains''  is not always well established, another practical setting is the translation of texts mixing several topics/domains. An additional requirement is to translate texts from domains unseen in training, based only on the unadapted system, which should then be made as strong as possible. 
% Residual adapters allow us to adapt the NMT model to any specific domain in a computationally cheap way. Could we adapt the NMT model to noisy text and topical text at once? While residual adapters are good at adapting to one specific domain, their performance still dramatically decreases for the other domains seen in training, as we show in our experiments. Therefore, we have to decide whether to use residual adapters manually, i.e, we have to know to which domain the text belongs a priori. Therefore, we would like to fuse a domain classifier to the architecture in order to weight the contribution of the adapters with respect to the topic-relatedness of the text.
\fyDone{Say differently: various implementations}

@@

\section{Residual adapters \label{sec:res}}
In this section, we describe the basic version of the residual adapter architectures \citep{houlsby19parameter, Bapna19simple}, as well as two original variants of this model.

@@

\subsubsection{The computation of adapter layers}
Our reference architecture is the Transformer model of \citet{Vaswani17attention}, which we assume contains a stack of layers both on the encoder and the decoder sides. Each layer contains two subparts, an attention layer, and a dense layer. Details vary from one implementation to another, we simply contend here that each layer $i \in \{1 \dots L\}$ (in the encoder or the decoder) computes a transform of a fixed-length sequence of $d$-dimensional input vectors $h^{i}$ into a sequence of output vectors $h^{i+1}$.

The residual adapter architecture extends this model by adding one new computation module for each layer. The $i^{\text{th}}$ residual adapter thus modifies $h^i$ by computing the following transformations:\fyDone{Use align env} \fyTodo{Add layer normalization, replace notation b by a}
\begin{align*}
  h^{i}_0 &= \mathbf{LN}(h^{i}) \\
  h^{i}_1 &= \mathbf{W}_{db}^{i}h_0^{i} + a^i_{1} \\
  h^{i}_2 &= \mathbf{ReLU}(h_1^{i})
  \\
%\end{align*}
%\begin{align*}
  h^{i}_3 &= \mathbf{W}_{bd}^{i}h_2^{i} + a^i_{2} \\
  \bar{h}^{i} &= h^{i}_3 + h^i.
\end{align*}
Overall, the  $i^{\text{th}}$ adapter is thus parameterized by matrices $\displaystyle{\mathbf{W}_{db}^{i}\in\mathbb{R}^{d\times b}}$,$\displaystyle{\mathbf{W}_{bd}^{i}\in\mathbb{R}^{b\times d}}$, bias vectors $\displaystyle{b^i_{1} \in \mathbb{R}^{b}}$, $\displaystyle{b^i_{2} \in \mathbb{R}^{d}}$, with $b$ the dimension of the adapter \fyTodo{($d \gg b$) not true}\fyDone{Check this}. For the sake of brevity, we will simply denote $h^{i}_3 = \operatorname{ADAP}^{(i)}(h^i)$, and $\theta_{\operatorname{ADAP}^{(i)}}$ the corresponding set of parameters.\fyDone{or is it $h_i$ ?}\fyDone{attention aux matrices $W_i$}

@@

\subsubsection{Design space and variants \label{sssec:design-space}}
This general architecture leaves open many design choices pertaining to the details of the network organization, the training procedure, and the corresponding objective function.

The first question concerns the number of adapter layers. While in principle, all Transformer layers can be subject to adaptation, it is nonetheless worthwhile to consider simpler adaptation schemes, which would only alter a limited number of layers. Such strategy might be especially relevant when the training data contains very small domains, as in the experiments of Section~\ref{sec:exp}, and for which a complete adaptation may not be necessary or/and or prone to overfitting. Likewise, it might be meaningful to explore ways to share subsets of adapters across domains. This in turn raises the issue of which layer(s) to adapt, a question that can be approached in the light of recent analyses of Transformers models, which conjecture that the higher layers encode global patterns with a more ``semantic'' interpretation, while the lower layers encode local patterns akin to morpho-syntactic information \cite{raganato18analysis}.

% The size of available corpora for each domain is extremely varying. For example, the corpus of domain \domain{tourism} in En-De contains only around $7\times10^3$ sentence pairs while the corpus of domain \domain{news} in En-De contains approximately $3\times10^6$ sentence pairs. For extremely small domains, we could economize the number of parameters by reducing the number of residual adapters in the architecture. While only a limited number of adapters are incorporated to the NMT model, which positions in the model are more important in the domain adaptation task? We would like to analyze the impact of position and number of residual adapters involved in the adapted model. It is conjectured that the higher layers represent more global patterns such as semantic while the lower layers represent more local patterns such as syntactic \cite{raganato18analysis}\fyTodo{Missing reference here}. The domain shift in local patterns and global patterns has not yet been studied. In this paper, we do not intend to study this aspect of the domain adaptation problem. We would like to give an extensive comparison between domain adaptation in different levels in the NMT model.

A related question concerns the regularization of adapter layers to mitigate overfitting. Reducing the number of adapters, or their dimensions, is simple, but such choices are difficult to optimize numerically -- an issue that becomes important as the number of domain grows. Less naive alternatives can also be entertained, such as applying weight decay or layer regularization to the adapter. Implementing these requires to modify the objective function in a way that still allows for a smooth optimization problem. For instance, weight decay applies a penalization on the weights of the adapters, complementing the cross-entropy term with a function of the norm of the parameters: \fyDone{The second summation also runs over $x,y$ ? I think not}
\begin{equation*}
  \begin{split}
    \bar{L} & = \frac{1}{\#(x,y)}\mathop{\sum}_{x,y}( - \log(P(y|x))) \\
    & + \lambda  \sum_{i \in \{1,..,6\} \otimes \{enc, dec\}} \normL{\theta_{\operatorname{ADAP}^{(i)}}}
  \end{split}
\end{equation*}
An alternative scheme is \emph{layer regularization}, which penalizes the output of the adapters, corresponding to the following objective:
\begin{equation*}
  \begin{split}
    \bar{L} & = \frac{1}{\#(x,y)}\mathop{\sum}_{x,y} (-\log(P(y|x)) \\
    & + \lambda \sum_{i \in \{1,..,6\} \otimes \{enc, dec\}} \normL{\operatorname{ADAP}^{(i)}(h_i(x,y))})
  \end{split}
\end{equation*}

Finally, another independent design choice relates to the training strategy for adapters. A first option is to generalize supervised domain adaptation to multi-domain adaptation and to proceed in two steps: (a) train a generic model with all the available data; (b) train each adapter layer with domain-specific data, keeping the generic model parameters unchanged. Another strategy is to adopt the view of \citet{Dredze08online}, where the multi-domain setting is viewed as an instance of multi-task learning \cite{Caruana97multitask} with each domain corresponding to a specific task. This suggests training all the parameters from scratch, as we would do in a multi-task mode. The generic parameters will still depend on all the available data, while each adapter will only be trained with the corresponding in-domain data.\fyFuture{Does everyone need a domain? Do we need this discussion - talk about sentence-level adaptation, better classifier pour MDL}

@@

Therefore, we would like to design a version of residual adapters that is more robust to such domain errors. This variant, called the \emph{gated residual adapter model}, relies on the training of a supplementary trained component that will help decide whether to activate, on a word per word basis, a given residual layer and to regulate the strength of this activation. To this end, we extend the highway version of residual adapters as follows.
\fyDone{Consistency of notations wrt section 2.1}

Formally, we replace the adapter computation of equation~\eqref{eq:highway-output} and take the adapted hidden (topmost) layer to be computed as (this is for domain $k$):
\begin{equation}
  \bar{h}^L = h^L + \displaystyle{\mathop{\sum}_{1 \leq i \leq L} \operatorname{ADAP}_k^i(h^i) \odot{} z_k(h^L)}, \label{eq:gated-output}
\end{equation}
where the scalar $z_k(h^L[t]) \in [0,1]$ measures the relatedness of the $t^{\text{th}}$ word $w_t$ to domain $k$. The more likely $w_t$ is in domain $k$, the larger $z_k(h^L[t])$ should be; conversely, for words\footnote{The term ``word'' is employed here by mere convenience, as systems only manipulate sub-lexical BPE units; furthermore, the values of the hidden representations $h^{i}$ at position $t$ depend upon all the other positions in the sentence.} that are not typical of any domain $k$ (eg.\ function words),  adaptation is minimum and the corresponding adapted encoder output ($\bar{h}^L[t]$) will remain close to the output of the generic model ($h^L[t]$). In our implementation, we incorporate two domain classifiers on top of the encoder and the decoder, that take the last hidden layer of the encoder (resp.\ decoder) as input and use the posterior probability $P(k|h^L[t])$ of domain $k$ as the value for $z_k(h^L[t])$.

Training gated residual adapters thus comprises three steps, instead of two for the baseline version:
\begin{enumerate}
\item learn a generic model with mixed corpora from multiple domains.
\item train a domain classifier on top of the encoder and decoder; during this step, the parameters of the generic model are frozen. This model computes the posterior domain probability $P(k|h^L[t])$ for each word $w_t$, based on the representation computed by the last layer.
\item train the parameters of adapters with in-domain data separately for each domain, while freezing all the other parameters.
\end{enumerate}
\fyFuture{is this classifier important, can we train with the rest of the system? Yes we could, but do we want the encoder to be good at classification?}\fyFuture{Check writing to indicate label smoothing for this has been used.}\fyFuture{Can we train a more complex training model ? should we use more monolingual data? etc}

@@

For all En$\rightarrow$Fr models, we set the embeddings size and the hidden layers size to~512. Transformers use multi-head attention with 8 heads in each of the 6 layers; the inner feedforward layer contains 2,048 cells. Residual adapters additionally use an adaptation block in each layer, composed of a 2-layer perceptron, with an inner ReLU activation function operating on normalized entries of dimension $b=1024$.\revision{ \citet{Bapna19simple} showed that the performance of adapted model increases with respect to the size of the inner dimension. In the experiments of \citet{Bapna19simple}, we note that $b=1024$, which is 2-folds larger than the dimension of Transformer layer, achieved performance closed to fine-tuned model. Therefore, we would like to apply this setting to our experiments.}
% The gated variant is made of a dense layer, followed by a layer normalization and a sigmoid activation.
% The domain control systems are exactly as their baseline counterparts (RNN and Transformer), with an additional 2 cells encoding the domain on the input layer.
Training uses a batch size of~12,288 tokens; optimization uses Adam with parameters $\beta_1=0.9$, $\beta_2= 0.98$ and Noam decay ($warmup\_steps=4,000$), and a dropout rate of $0.1$ for all layers.\revision{ For \system{mixed} model, we use initial learning 1.0 and the concatenation of the validation sets of 6 domains as the validation set of \system{mixed}. In the fine-tuning experiments, we continue training using \system{mixed} as starting point, we use the same learning rate schedule, continuing the incrementation of the number of steps. In the multi-task training, we use the same learning rate schedule as \system{mixed}, for each iteration, we sample a domain from the pool of 6 domains with the probability proportional to the size of training data of each domain, then we sample a batch of~12,288 tokens from training data of the picked domain, the shared parameters and the parameters of the corresponding adapter are updated by gradients.}\fyDone{Describe the block adaptation layer - voir slides}

@@

\iffalse{
\begin{table*}[htbp]
  \centering
  \begin{tabular}{|l|l|} \hline
    Model & params \\ \hline 
    \system{Transformer-En-Fr}  & 65 \\
    \system{Residual Adapter-En-Fr} & 1 \\
    \system{Transformer-En-De}  & 213 \\
    \system{Residual Adapter-En-De} & 4 \\
     \hline
  \end{tabular}
  \caption{Number of parameters ($\times 10^6$)}
  \label{tab:params}
\end{table*}
}
\fi

@@

By varying the number and position of residual adapters (see Section~\ref{ssec:architecture}), we then contrast several implementations of residual adapters. \fyDone{Fix style here} Because the set of possible configurations is large, we only perform experiments for layers $i= 2, 4, 6$ (both for the encoder and decoder). Two settings are considered: keeping just one adapter or keeping the three. The trend is the same for the two language directions: suppressing adapters always hurts the overall performance, albeit by a small margin: having six adapters is better than three, which is better than keeping only one. With only one adapter active, we observe small, insignificant changes in performance when varying the adapter's depth. \revision{Based on these experiments, it seems that adaptation remains useful even for the deeper layers of the architecture.}\fyFuture{More configurations ?}

@@

Results in Tables~\ref{tab:performance-en-fr-pos-reg} and \ref{tab:performance-en-de-pos-reg} show that regularizing the adapter model can positively affect the test performance for the smallest domains (this is especially clear for weight-decay (\system{Res-Adap-WD}) in En$\rightarrow$De), at the cost however of a small drop in performance for the other domains. Using layer regularization proves here to be comparatively less effective. Finding better ways to set the regularization parameters, for instance by varying $\lambda$ for each domain based on the available supervision data, is left for future work.  
\fyDone{How is the weight decay parameter set ?}
\fyDone{Why is regularization not helping? It helps for the small domain - domain-specific regularization ??}
